---
title: "Races or Tournaments? Theory and Evidence in the Field"
subtitle: "Appendices"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
author: "Contact: Andrea Blasco (ablasco@fas.harvard.edu)"

output:
  html_document:
    theme: cosmo
    code_folding: hide
    highlight: kate #monochrome
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
# output:
#   pdf_document:
#     highlight: kate #monochrome
#     toc: true
#     template: ~/Templates/Pandoc/boyd.latex
#     keep_tex: false
# header-includes: 
#   - \usepackage{titlesec}
#   - \newcommand{\sectionbreak}{\clearpage}
#   - \input{defs.tex}
#   - \setlength\parindent{0pt}

...

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	cache=FALSE,
	echo=TRUE,
	message=FALSE,
	warning=FALSE,
	error=TRUE,
	tidy=TRUE,
	tidy.opts=list(width.cutoff=60),
	cache.path="Cache/",
	fig.path="Figures/")

child_docs <- setdiff(list.files(".", pattern='*.Rmd'), "_note.Rmd")
```

```{r child = child_docs}
```
 