# Identification (old)

## Example with the Uniform distribution

Imagine that the distribution $F_\theta$ is uniform on $(0, \theta)$ and the cost ability $c_a(x) = 1/x$. Then the zero profit condition \eqref{zero profit} becomes

\begin{align}
	\left(\frac{m}{\theta}\right)^{n-1} m = \beta x_l 
\end{align}

Solving for the marginal type gives:

\begin{align}
	m = \left(\beta x_l  \theta^{n-1}\right)^{1/n}
\end{align}

So, the probability of participation \eqref{probability} becomes

\begin{align}
	p(x_l, \theta) 
		& = 1 - F_\theta(m) \\
		& = 1 - \frac{\left(\beta x_l \theta^{n-1}\right)^{1/n} }{\theta} \\
		& = 1 - \left(\frac{\beta x_l}{\theta}\right)^{1/n}.
\end{align}

Because it is non-linear in its parameters, the model is not identifiable [need proof]. However, we can re-parametrize the model with $\beta/\theta \equiv \eta$ to make the parameter $\eta$ identifiable. 

```{r, echo=FALSE, fig.cap="Simulated distribution of participants with uniform distribution. Parameters are $\\eta=0.1$ and $n=15$. Data simulated 500 times. Dashed curve had higher costs ($x_l=1.5$) than the solid curve ($x_l=0.5$)."}

# Plot curves
n <- 15
x1 <- 0.5
x2 <- 1.5
beta <- 0.1
p <- function(x, beta, theta, n) {
	1 - (beta * x / theta)^{1/n}
}
rooms <- 500
rooms.x1 <- replicate(rooms, rbinom(n=1, size=n, p=p(x1, beta, 1, n)))
rooms.x2 <- replicate(rooms, rbinom(n=1, size=n, p=p(x2, beta, 1, n)))

rooms.x1.prob <- data.frame(table(rooms.x1)/rooms)
rooms.x2.prob <- data.frame(table(rooms.x2)/rooms)

par(mar=c(4,4,0,0))
plot(NA, NA, ylim=range(rooms.x1.prob[,2], rooms.x2.prob[,2])
	, xlim=c(0,10), ylab="probability", xlab="Room participants")
lines(rooms.x1.prob, type='b')
lines(rooms.x2.prob, type='b', col=2, pch=2, lty=2)
```


## Example with Beta distribution

In this example, we assume abilities are drawn from a lognormal distribution with parameters $\mu$ and $\sigma$. 

## Estimation

Our aim is to estimate the parameters $\theta$ and $\beta$ and to evaluate whether costs are different between the three competition modes under study: race, tournament, tournament + requirement. 

It is natural to estimate the parameters $\theta$ by maximum likelihood because the ability distribution (and therefore the distribution of the outcomes) is known. The estimation criterion used here is the maximization of the _deviance_.

The deviance is:
\begin{equation}
	D(\theta) = -2 \sum Y \log (\frac{N p(x, \theta)}{Y}) 
		+ (N - Y) \log ( \frac{N - N p}{N - Y}).
\end{equation}

Note that the deviance is a function of the likelihood (see xxx pg. xxx). 

```{r}

# Define likelihood
structreg <- function(x, y, wt = rep(1, length(y)), intercept = TRUE, start = rep(0, p), ...) {
	fmin <- function(beta, X, y, w) {
		p <- 1 - (beta %*% X)^(1/n)  # Function of parameters
		-sum(2 * w * ifelse(y, log(p), log(1-p))) 
	}
# 	gmin <- function(beta, X, y, w) {
# 		# Gradient here!
# 	}
	if(is.null(dim(x))) dim(x) <- c(length(x), 1)
	dn <- dimnames(x)[[2]]
	if(!length(dn)) dn <- paste("Var", 1:ncol(x), sep="")
	p <- ncol(x) + intercept
	if(intercept) {x <- cbind(1, x); dn <- c("(Intercept)", dn)} 
	if(is.factor(y)) y <- (unclass(y) != 1)
#	fit <- optim(start, fmin, gmin, X = x, y = y, w = wt, method = "BFGS", ...)
	fit <- optim(0.5, fmin, X = x, y = y, w = wt, method = "BFGS", ...)
	names(fit$par) <- dn
	cat("\nCoefficients:\n"); print(fit$par)
	cat("\nResidual Deviance:", format(fit$value), "\n") 
	cat("\nConvergence message:", fit$convergence, "\n")
	invisible(fit)
}

# Create data 
ilogit <- function(x) exp(x)/(1+exp(x))
n <- 50000
competitors <- 5
x1 <- rchisq(competitors, df=3)
x2 <- runif(n)
x3 <- rnorm(n)
X <- cbind(rep(1, length(x1)), x1, x2, x3)
betas <-  c(-2.5, 0.1, 0.25, 0.5)
fc <- ilogit(X %*% betas) ## This the fixed cost
p <-  1 - fc^(1/competitors) 
y <- rbinom(n=n, size=competitors, prob=p)

# Objective function
fmin <- function(beta, X, y, w, competitors) {
	p <- 1 - ilogit(X %*% beta)^(1/competitors)  # Function of parameters
	-sum(2 * w * ifelse(y, log(p), log(1-p))) 
}

b.start <- runif(length(betas))
(out <- optim(b.start, fmin, X=X, y=y, competitors=competitors, w=rep(1, length(y)))$par)

rbind(out/competitors, betas)

```

## More examples

Suppose F is lognormal($\mu, \sigma$), then the zero profit condition is:

\begin{align}
	\left[\frac 12 +\frac 12 \left(\frac{\log{m}-\mu}{\sqrt{2\sigma}}\right)\right] = a^{-1/(n-1)} K_l
\end{align}

```
# xxx
zeroprofit <- function(x, mu, sigma, n) {
	x * plnorm(x, meanlog=mu, sdlog=sigma)^(n-1)
}
for (i in 3:10)
	curve(zeroprofit(x, i, 10, 10), add=TRUE, lty=i)
```



```{r}
# Clear
rm(list=ls())

# Libraries 
require(xtable)
require(moments)
require(races)

# Data
data(races)
attach(races)

# TABLE 1. 
size <- ave(submit, paste(treatment, room), FUN=length)
(m <- aggregate(submit ~ room + size + treatment, FUN=sum))
```

 