---
title: Contest package in R
author: Andrea Blasco
institute: ablasco@fas.harvard.edu
date: \today
output: 
  beamer_presentation:
    slide_level: 2
    template: ~/Templates/Pandoc/boyd.beamer
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

# What does the `contest` package do?

## What does the `contest` package do?

* This package **solves** an extension of the [Moldovanu and Sela (2001)](https://www.aeaweb.org/articles?id=10.1257/aer.91.3.542)'s contest model that encompasses race contests (the first wins) and tournament contests (the best wins).

* It provides an easy way to **compute equilibria** using numerical integration when the model does not allow for explicit solutions (which happens many times)

* Helps running **Monte Carlo studies** to find optimal contest design (number of competitors, prize structure, deadlines, entry requirements, etc.)

* Facilitates **estimation** of model's parameter from real data

## Installation

The latest version of the package is on GitHub. 

First, you need to install and invoke the `devtools` package.

```r
install.packages('devtools')
library(devtools)
```

Second, you need to install the package from the repo `"mrblasco/contest"` with the following command 

```r
install_github("mrblasco/contest")
```


## The MS (2001)'s contest model

* It is an "all-pay" contest model: everyone incurs a cost from effort that is paid before uncertainty about rivals' actions is fully realized

* Players compete for $k=1,..., q$ prizes of decreasing value $v_1\geq v_2\geq ...\geq v_q$ with $\sum v_k = 1$.

* Decide simultaneously completion time $t$ and performance $y$ (or equivalently: speed $s$ and performance $y$)

* Players have ability $a\sim F$ on a bounded interval.

* Costs are multiplicative: $c(a, y, t)$ with elasticities $\alpha<0$, $\beta>0$, and $\gamma<0$.  


## The `contest()` function

Arguments:

* `x`: a vector of realized abilities
* `n`: the number of competitors in the contest
* `type`: either a race or a tournament
* `prize`: a vector with prizes for 1st and 2nd place (prizes higher than 2nd place are ignored)
* `elasticity`: a vector with cost elasticities
* `p, d`: distribution and density functions for abilities
* `xlim`: interval for ability distribution
* `deadline, target`: parameters of the contest


## Outcomes

Computes the symmetric equilibrium via `R` builtin numerical integration  tool `integrate()`.

Values:

* `ability`: a vector of input abilities
* `score`: a vector of equilibrium performances
* `timing`: a vector of equilibrium completion times
* `utility`: a vector of expected payoffs
* `marginal.type`: ability level below which it is not optimal to enter the contest
* `params:` list of contest parameters

## Usage
 
```r
library(contest)
ability <- runif(10) # uniformly distributed (default)
contest(
	sort(ability)
	, n=10 # 10 players	
	, prize=c(1,0) # Winner-takes-all
	, deadline=10 # Time upper-limit
	, target=0.5 # Performance lower-limit
)	
```

## Example: compare ability distributions

```{r, echo=FALSE}
par(mfrow=c(2, 4))
x <- seq(0, 1, length=50)

# Uniform
curve(dunif, xlab="Ability ~ Uniform", ylab="density")
title("Ability distribution")
plot(contest(x, n=4, p=punif,d=dunif), pch=16, cex=.5)

# Beta
curve(dbeta(x, shape1=2, shape2=5), xlab="Ability ~ Beta(2, 5)", ylab="density")
title("Ability distribution")
plot(contest(x, n=4, p=pbeta,d=dbeta, shape1=2,shape2=5), pch=16, cex=.5)
```


## Example: compare races and tournaments

```{r, echo=FALSE}
library(contest)
x <- seq(0, 1, length=50)
race <- contest(x, n=4, type='race', prize=c(1,0), deadline=10, target=0.5
		, p=pbeta,d=dbeta, shape1=2,shape2=5, elasticity=c(-1,2,-0.5))
tournament <- contest(x, n=4, type='tournament', prize=c(1,0), deadline=10, target=0.5
		, p=pbeta,d=dbeta, shape1=2,shape2=5, elasticity=c(-1,2,-0.5))

par(mfrow=c(2, 3))
plot(tournament, pch=16, cex=.5)
plot(race, pch=16, cex=.5)
```

# Monte Carlo Experiments
## Monte Carlo Method

- Random sampling from true distribution to compute expected values (e.g., expected average performance)

- Let $y^*(\cdot)$ $t^*(\cdot)$ denote the equilibrium functions for performance and time

- The expected average (equilibrium) performance under $F$ is

$$
	E_F[y*] = \int_{0}^1 y^*(x) dF(x)
$$

- Let $a=(a_1, a_2, ..., a_s)$ be $s$ realization of abilities under $F$

- By the law of large numbers, the sample mean computed on several replica of $a$ converges (in distribution) to the expected value


## Example: expected top performance

In equilibrium, the expected top performance under $F$ is

$$
	\theta = E_{F}[\max\{y_1^*,y_2^*,...,y_n^*\}] = E_{F_{(n)}}[y^*]
$$

where $F_{(n)}$ is the distribution of the $n$-th order statistics of the $y$'s.

Monte Carlo estimation:

- Simulate $s$ abilities $a_1, a_2,...,a_s$ from $F_{(n)}$

$$
	\hat\theta = \frac{1}{s}\sum y^*(a_i) 
$$


## Example: expected top performance in R

```{r}
mc.contest <- function(R, n=10, ...) {
	a <- runif(R*n) # abilities
	a.max <- tapply(a, rep(1:R, each=n), max)
	scores <- contest(a.max, n=n, ...)$score
	mean(scores) # Estimated expected top performance
}
mc.contest(R=100, type='tournament')
``` 

## How many replications?

```{r replications, echo=FALSE}
nsim <- seq(50, 1000, by=100)
out <- sapply(nsim, function(x) replicate(50, mc.contest(R=x, type='tournament')))
colnames(out) <- nsim
boxplot(out, xlab="Monte Carlo replications", ylab="Estimates")
```

# Optimal contest design
## Contest designer's revenues

* When timing is not an issue, revenues depend only upon performance.

* Imagine the contest designer wants to maximize expected top performance:

$$
	\pi_{cd} = E_{F_{(n)}}[y^*] 
$$

* What is the **optimal allocation of prizes**? 
* What is the **optimal target**?
* What is optimal between **race and tournament competitions**?

## The Optimal Allocation of Prizes in a Tournament

MS (2001) proves that:

> [ With **linear** or **concave** cost functions,] "it is optimal to allocate the entire prize sum to a single first prize. [...] With **convex** cost functions, the beneficial effect of the second prize on middle- and low- ability players is amplified, while the advantage of having one prize (which strongly motivates high-ability contestants) is decreased."

## The optimal allocation of prizes in R

```{r optim-prize}
# Linear costs, e.g., x^1
f <- function(x, elasticity) {
	mc.contest(R=1e4, n=10, prize=c(x, 1-x), elasticity=elasticity)
}
set.seed(1)
optimize(f, interval=c(0.5, 1), maximum=TRUE, elasticity=c(-1, 1, -1))$maximum

# Convex costs, e.g., x^2
set.seed(1)
optimize(f, interval=c(0.5, 1), maximum=TRUE, elasticity=c(-1, 2, -1))$maximum
```

## Optimal Target

Blasco, Boudreau, Lakhani and Menietti (2017)'s paper on races vs tournaments shows that:

1. Optimal target is always lower in a tournament compared to a race.

2. Tournaments generally dominate races in terms of (ex-ante) top performance

3. Difference between tournaments and races reduces when the distribution is asymmetric

##  Optimal Target in a Tournament

```{r optim-tournament}
f <- function(x, n=10, R=350, ...) {
	mc.contest(R=R, n=n, target=x, ...)
}
# optimal target
set.seed(1)
optimize(f, type='tournament', interval=c(0, 1), maximum=TRUE) 
```

## Optimal Target in a Race

```{r optim-race}
set.seed(1)
optimize(f, type='race', interval=c(0, 1), maximum=TRUE) 
```

## Races vs tournaments

```{r racs-vs-tournaments, echo=FALSE}
m <- numeric(2)
m[1] <- optimize(f, type='tournament', interval=c(0, 1), maximum=TRUE
				, p=pbeta, d=dbeta, shape1=2, shape2=4)$maximum 
m[2] <- optimize(f, type='race', interval=c(0, 1), maximum=TRUE
				, p=pbeta, d=dbeta, shape1=2, shape2=4)$maximum

x <- seq(0.6, 1, length=200)
par(mfrow=c(2, 3))

# Uniform distribution
tournament <- contest(x, n=10, type='tournament', target=0.189)
race <- contest(x, n=10, type='race', target=0.49)

curve(dunif, xlab="Ability ~ Uniform", ylab="density")
title("Ability distribution")
plot(x, tournament$score, pch=16, cex=.5, type='b', xlab="ability", ylab="Score")
points(x, race$score, pch=16, cex=.5, type='b', col=2)
plot(x, tournament$timing, pch=16, cex=.5, type='b', xlab="ability", ylab="Timing")
points(x, race$timing, pch=16, cex=.5, type='b', col=2)

## Beta distribution
tournament2 <- contest(x, n=10, type='tournament', target=m[1], p=pbeta, d=dbeta, shape1=2, shape2=4)
race2 <- contest(x, n=10, type='race', target=m[2], p=pbeta, d=dbeta, shape1=2, shape2=4)

curve(dbeta(x, 2, 4))
title("Ability distribution")
plot(x, tournament2$score, pch=16, cex=.5, type='b', xlab="ability", ylab="Score")
points(x, race2$score, pch=16, cex=.5, type='b', col=2)
plot(x, tournament2$timing, pch=16, cex=.5, type='b', xlab="ability", ylab="Timing")
points(x, race2$timing, pch=16, cex=.5, type='b', col=2)

```

## Suppose now that time matters ... 

* Imagine the contest designer wants to maximize: 

$$
	\pi_{cd} = E_{F_{(n)}}[y^* + \tau t^*] 
$$

where the intensity of completion-time preferences are denoted by $\tau \leq 0$. 

* What is the optimal prize structure? 
* What is the optimal target?
* What is optimal between a race and tournament competition?

## Update the contest simulator

```{r}
mc.contest2 <- function(R, n=10, tau, ...) {
	a <- runif(R*n) # abilities
	a.max <- tapply(a, rep(1:R, each=n), max)
	out <- contest(a.max, n=n, ...)
	scores <- contest(a.max, n=n, ...)$score
	timing  <- contest(a.max, n=n, ...)$timing
	mean(scores + tau*timing)
}
mc.contest2(R=300, tau=-0.1)
``` 

## The optimal allocation of prizes (with time) in R

```{r optim-time}
# Linear costs, e.g., x^1
f <- function(x, tau, elasticity, R=1e3, n=10) {
	mc.contest2(R=R, n=n, prize=c(x, 1-x),  tau=tau, elasticity=elasticity)
}
set.seed(1)
optimize(f, interval=c(0.5, 1), maximum=TRUE, tau=-0.1, elasticity=c(-1, 1, -1))$maximum

# Convex costs, e.g., x^2
set.seed(1)
optimize(f, interval=c(0.5, 1), maximum=TRUE, tau=-0.1, elasticity=c(-1, 2, -1))$maximum
```

## The optimal target (with time)

BBLM (2017)'s paper shows that:

1. There exist a threshold in $\tau$ above which races are preferred to tournaments, and viceversa


## The optimal target (with time) in R

```{r time-target}
f <- function(x, tau, elasticity, R=1e3, n=10, ...) {
	mc.contest2(R=R, n=n, target=x,  tau=tau, ...)
}

# Optimal target in a tournament
set.seed(1)
optimize(f, tau=-0.6, type='tournament', interval=c(0, 1), maximum=TRUE, p=pbeta, d=dbeta,shape1=2,shape2=4) 

# Optimal target in a race
set.seed(1)
optimize(f, tau=-0.6, type='race', interval=c(0, 1), maximum=TRUE, p=pbeta, d=dbeta,shape1=2,shape2=4) 

```


# Estimation -- Approach #1
## Estimating the ability distribution 

* The ability distribution $F(\cdot \mid \theta)$ is known up to a paramter $\theta$ that we want to estimate from the data

* We observe a sample of $y_1,...,y_s$ from $F_{Y}(\cdot \mid \theta)$ where $y_i$ is observed only when it is higher than a target $y_0$, otherwise is zero.

* At equilibrium, we have that $Y= y^*(A; \theta)$ where $y^*(\cdot;\theta)$ is a monotonic increasing function with inverse denoted by $\phi=y^{*-1}$. 

## The likelihood function

* Let $\mathbf{1}_{y_0}(x)$ denote an indicator function that is 1 when $x>y_0$, and 0 otherwise

$$
	\mathcal{L}(\theta) = \prod_{i=1}^s F_{Y}(y_0\mid\theta)^{1-\mathbf{1}_{y_0}(y_i)}f_{Y}(y_i\mid\theta)^{\mathbf{1}_{y_0}(y_i)}
$$

The log likelihood is

\begin{align*}
	\log\mathcal{L}(\theta) & = \sum_{i=1}^s [1-\mathbf{1}_{y_0}(y_i)] \log F_{Y}(y_0) 
										+ \mathbf{1}_{y_0}(y_i) \log f_{Y}(y_i)\\
			 & = n_0 \log F_{Y}(y_0) +  \sum_{i=1}^s \mathbf{1}_{y_0}(y_i) \log f_{Y}(y_i)
\end{align*}

## The likelihood function, cont'd

We substitute $F_{Y}$ with the known ability distribution $F$ (change of variable)
\begin{align*}
	\log\mathcal{L}(\theta) & = n0 \log F(\phi(y_0;\theta)\mid\theta) 
					+ \sum_{i=1}^s \mathbf{1}_{y_0}(y_i) \log f(\phi(y_i;\theta)\mid\theta) \phi^\prime(y_i;\theta))
\end{align*}
where $\phi(y_i;\theta)$ represents the theoretical ability $a_i(\theta)$ of an individual with a score $y_i$ conditional on the ability distribution's parameter $\theta$


## Compute the inverse $\phi$ and its first derivative

The invertible function is: $y^* = (y_0^{\beta} + \sum v_k \int g_k(x) dx )^{1/\beta}$

We use **cubic spline interpolation** to compute $\hat\phi(\cdot;\theta)$ and $\hat\phi^\prime(\cdot;\theta)$.


```{r spline}
# First compute contest equilibrium function on a grid under a given theta
x.100 <- seq(0, 1, length=100)
f <- function(x) contest(x, n =4, p=pbeta,d=dbeta,shape1=2,shape2=4)$score
y <- f(x.100)

# Then we compute the interpolated bid function
fun <- splinefun(x.100, y)
# or the inverse
fun.inv <- splinefun(y, x.100)
```


## Example of spline interpolation

```{r spline-example, echo=FALSE}
# Plot results
curve(f, from=0, to=1)
curve(fun, from=0, to=1, add=TRUE, lty=2, col=2)
```

## Define $\log \mathcal{L}$ in R

```{r loglik}
loglik <- function(par, data)  {
	y <- data$y; n <- data$n # parse data
	x.grid <- seq(0, 1, length=100) # Evaluation "grid"
	out <- contest(x.grid, n=n, type="tournament", p=pbeta, d=dbeta
					, shape1=par[1], shape2=par[2])
	entrants <- y>0
	n0 <- sum(!entrants)
	a0 <- out$marginal.type # Marginal type given theta
	inv <- splinefun(x=out$score, y=x.grid)
	a <- inv(y[entrants])
	a <- ifelse(a>1, 1, a) # Trim at the top
	a.deriv <- inv(y[entrants], deriv=1)
	ll <- n0 * pbeta(a0, shape1=par[1], shape2=par[2], log.p=TRUE) + 
	sum(dbeta(a, shape1=par[1], shape2=par[2], log=TRUE) * a.deriv)
	return(ll)
}
```

## Check $L(\tilde\theta)\leq L(\theta)$ for any $\tilde\theta\neq\theta$

Assume we observe $m$ replications of contests with $j$ competitors

```{r lokglik-sim, echo=FALSE}
# Data simulator
sim <- function(m, j, ...) {
	args <- list(...)
	x <- rbeta(m * j, args$shape1, args$shape2)
	out <- contest(x, n=j, p=pbeta, d=dbeta, ...)
	return(list(y=out$score, n=j, x=x))
}

# Simulate data
# set.seed(1)
# dat <- sim(m=100, j=4, type='tournament', shape1=2, shape2=4)
# plot(dat$x, dat$y, cex=.5, pch=16, ylim=c(0, 1), xlim=c(0, 1))
# curve(contest(x, n=4, p=pbeta,d=dbeta,shape1=2,shape2=4)$score, add=TRUE)
# curve(contest(x, n=4, p=pbeta,d=dbeta,shape1=2,shape2=4.5)$score, add=TRUE, lty=2)
# curve(contest(x, n=4, p=pbeta,d=dbeta,shape1=2,shape2=3.5)$score, add=TRUE, lty=3)

# Compute ll
# par.grid <- pretty(seq(1.01, 4.5, length=10))
# ll <- replicate(100, {
# 	dat <- sim(m=10, j=4, shape1=2, shape2=4)
# 	sapply(par.grid, function(x) loglik(c(2, x), dat))
# })
# 
# boxplot(t(ll[-c(8), ]), names=par.grid[-8], outline=FALSE)
# 
# plot(par.grid, ll, ylab="Log Likelihood", xlab='theta', pch=16, cex=.5)
# abline(v=5)

set.seed(1)
theta <- 5
dat <- sim(m=30, j=4, shape1=2, shape2=theta)
par.grid <- seq(1.01, 8, length=50)
ll <- sapply(par.grid, function(x) loglik(c(2, x), dat))
plot(par.grid, ll, ylab="Log Likelihood", xlab='theta', pch=16, cex=.5)
abline(v=theta, lty=3, col=2)
```

## Open Issues

- Understanding sources of bias 
	+ None seen so far
- Double checking implementation 
- Work out simple example by hand


# Estimation -- Approach #2
## Estimating ability distribution from entry data

Under the model:  $\text{entry} \iff \text{ability}\geq a_0$


Assume: 

$$
	\text{ability}_i - a_{0} = \underbrace{\sum g_i(X_i\beta) + \epsilon_i}_{\text{index}}
$$

with $\epsilon \sim F(\cdot\mid\theta)$ which is known up to a parameter $\theta$

Potential limitations:

1. $a_0$ might depend on $\theta$ and $X\beta$ (we ignore such dependence)
2. Competitors might observe $X$'s (violating "ex-ante identical" assumption)


## Estimating the entry model from experimental data

```{r, echo=FALSE}
races$hours <- with(races, week1 + week2 + week3 + week4)
races$hour.imp <- impute(races$hours, "random")
summary(fit <- glm(submit ~ treatment*log(mm_reg)*mm_events*hour.imp + I(!is.na(paid)), data=races, family=binomial))

# plot
yhat <- predict(fit)
plot(yhat, jitter(fit$y,.2), pch=16, cex=.5, col=ifelse(yhat>0,"brown","blue"), xlab="Ability ~ logistic", ylab="Entry (jittered)", yaxt='n')
axis(2, at=c(0,1))
curve(plogis, add=TRUE)
```

## Simulate ability distribution

```{r, echo=FALSE}
simulate.ability <- function(fit) {
	index <- predict(fit)
	rlogis(length(index), location=index, scale=1)
}
hist(replicate(100, simulate.ability(fit)), "Scott")
```


## Contest simulations

```r
kde <- function(x, ...) approxfun(density(simulate.ability(fit)))
kde.cum <- function(x, ...) integrate(kde, lower=0, upper=x)$value
contest(seq(0, 1), n=15, d=kde, p=kde.cum)
```

## Results

```{r, echo=FALSE}
par(mfrow=c(2, 3))
plot(contest(seq(0, 1, length=10), n=15, d=dbeta, p=pbeta, shape1=2,shape2=3, type='tournament'))
plot(contest(seq(0, 1, length=10), n=15, d=dbeta, p=pbeta, shape1=2,shape2=3, type='race'))
```