---
title: Contest package in R
author: Andrea Blasco
institute: ablasco@fas.harvard.edu
date: \today
output: 
  beamer_presentation:
    slide_level: 2
    template: ~/Templates/Pandoc/boyd.beamer
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

# Monte Carlo Experiments
## Monte Carlo Method

- Random sampling from true distribution to compute expected values (e.g., expected average performance)

- Let $y^*(\cdot)$ $t^*(\cdot)$ denote the equilibrium functions for performance and time

- The expected average (equilibrium) performance under $F$ is

$$
	E_F[y*] = \int_{0}^1 y^*(x) dF(x)
$$

- By the law of large numbers, the sample mean computed on several replications of this process converges (in distribution) to the expected value


## Example: expected top performance

In equilibrium, the expected top performance under $F$ is

$$
	\theta = E_{F}[\max\{y_1^*,y_2^*,...,y_n^*\}] = E_{F_{(n)}}[y^*]
$$

where $F_{(n)}$ is the distribution of the $n$-th order statistics of the $y$'s.

Monte Carlo estimation:

- Simulate $s$ abilities $a_1, a_2,...,a_s$ from $F_{(n)}$

$$
	\hat\theta = \frac{1}{s}\sum y^*(a_i) 
$$


## Example: expected top performance in R

```{r}
## Monte Carlo simulator function
mc.contest <- function(R, n=10, ...) {
	a <- runif(R*n) # abilities
	a.max <- tapply(a, rep(1:R, each=n), max)
	scores <- contest(a.max, n=n, ...)$score
	mean(scores) # Estimated expected top performance
}
mc.contest(R=100, type='tournament')
``` 

## How many replications?

```{r replications, echo=FALSE, fig.cap="Each boxplot shows 50 monte carlo simulations with the number of replications along the x-axis"}
nsim <- c(seq(50, 1000, by=100))
out <- sapply(nsim, function(x) replicate(50, mc.contest(R=x, type='tournament')))
colnames(out) <- nsim
boxplot(out, xlab="Monte Carlo replications", ylab="Estimates")
```

## Example: optimal prize allocation

```{r optimal-prize, echo=FALSE, fig.cap="Each boxplot shows 50 monte carlo simulations with the number of replications along the x-axis"}
# Function to optimize
f <- function(x, R, ...) mc.contest(R, n=10, prize=c(x, 1-x), ...)
nsim <- c(100, 300, 500, 1000)
out <- sapply(nsim, function(x) replicate(50, optimize(f, R=x, interval=c(0.5, 1), maximum=TRUE)$maximum))
colnames(out) <- nsim
boxplot(out, xlab="Monte Carlo replications", ylab="Optimal prize allocation")
```

# Estimation
## Estimating an unknown ability distribution 

* The ability distribution $F(\cdot \mid \theta)$ is known up to a paramter $\theta$ that we want to estimate from the data

* We observe a sample from $Y_1, Y_2, ..., Y_s$ iid from $F_{Y}(\cdot \mid \theta)$ where $y_i$ is observed only when it is higher than a target $y_0$, otherwise is zero.

* At equilibrium, there's a structural relationship between performance and abilities $Y= y^*(A; \theta)$ where $y^*(\cdot;\theta)$ is a monotonic increasing function with inverse denoted by $\phi=y^{*-1}$. 

## The likelihood function

* Let $\mathbf{1}_{y_0}(x)$ denote an indicator function that is 1 when $x>y_0$, and 0 otherwise

$$
	\mathcal{L}(\theta) = \prod_{i=1}^s F_{Y}(y_0\mid\theta)^{1-\mathbf{1}_{y_0}(y_i)}
	f_{Y}(y_i\mid\theta)^{\mathbf{1}_{y_0}(y_i)}
$$

The log likelihood is

\begin{align*}
	\log\mathcal{L}(\theta) & = \sum_{i=1}^s [1-\mathbf{1}_{y_0}(y_i)] \log F_{Y}(y_0) 
										+ \mathbf{1}_{y_0}(y_i) \log f_{Y}(y_i)\\
			 & = n_0 \log F_{Y}(y_0) +  \sum_{i=1}^s \mathbf{1}_{y_0}(y_i) \log f_{Y}(y_i)
\end{align*}

## The likelihood function, cont'd

We don't know  $F_{Y}$ but we the known ability distribution $F$. 

With a change of variable, we obtain
\begin{align*}
	\log\mathcal{L}(\theta) & = n0 \log F(\phi(y_0;\theta)\mid\theta) 
					+ \sum_{i=1}^s \mathbf{1}_{y_0}(y_i) \log f(\phi(y_i;\theta)\mid\theta) \phi^\prime(y_i;\theta))
\end{align*}
where $\phi(y_i;\theta)=a_i(\theta)$ represents the theoretical ability of player $i$ with a score $y_i$ given the distribution's parameter $\theta$. 


## Compute the inverse $\phi$ and its first derivative

The invertible function is: $y^* = (y_0^{\beta} + \sum v_k \int g_k(x) dx )^{1/\beta}$

We use **cubic spline interpolation** to compute $\hat\phi(\cdot;\theta)$ and $\hat\phi^\prime(\cdot;\theta)$.


```{r spline}
# First compute contest equilibrium function on a grid under a given theta
x.100 <- seq(0, 1, length=100)
f <- function(x) contest(x, n=4, p=pbeta, d=dbeta, shape1=2, shape2=4)$score
y <- f(x.100)

# Then we compute the interpolated bid function
fun <- splinefun(x.100, y)
fun.inv <- splinefun(y, x.100) # or the inverse
```


## Example of spline interpolation

```{r spline-example, echo=FALSE}
# Plot results
curve(f, from=0, to=1)
curve(fun, from=0, to=1, add=TRUE, lty=2, col=2)
```

## One-parameter example

```{r loglik}
loglik <- function(param, data)  {
	# parse data
	S1 <- 2
	S2 <- param
	y <- data$score
	n <- data$params$n
	target <- data$params$target
	deadline <- data$params$deadline
	# Simulate the model given params
	x.grid <- seq(0.0001, 1, length=100) # Evaluation "grid"
	out <- contest(x.grid, n=n, type="tournament", p=pbeta, d=dbeta
					, shape1=S1, shape2=S2, target=target, deadline=deadline)
	inv <- splinefun(x=out$score, y=x.grid)
	a <- inv(y)
	a <- ifelse(a>1, 1, a) # Trim at the top
	a <- ifelse(a<=0, 0.000001, a) # Trim at the bottom
	a.deriv <- inv(y, deriv=1)
	ll <- mean(dbeta(a, shape1=S1, shape2=S2, log=TRUE) * a.deriv)
	return(ll)
}
```

## Check $L(\tilde\theta)\leq L(\theta)$ for any $\tilde\theta\neq\theta$

Assume we observe $m$ replications of contests with $j$ competitors

```{r lokglik-sim, echo=FALSE}
# Data simulator
sim <- function(m, j, ...) {
	args <- list(...)
	x <- rbeta(m * j, args$shape1, args$shape2)
	out <- contest(x, n=j, p=pbeta, d=dbeta, ...)
	return(out)
}

dat <- sim(10, j=10, shape1=2, shape2=4, target=0.01)

# Check feasibility
f <- function(x, maxvalue) {
	contest(1, n=10, target=0.01, p=pbeta, d=dbeta, shape1=2, shape2=x)$score - maxvalue
}
uniroot(f, maxvalue=max(dat$score), interval=c(1, 10))$root

optimize(loglik, data=dat, interval=c(1, 5.33), maximum=TRUE)

p.grid <- seq(2, 4.5, length=50)
out <- sapply(p.grid, function(x)loglik(param=x, data=dat))
plot(p.grid, out)

