---
title: "Races or Tournaments?"
author: 
    - \textbf{Andrea Blasco}\footnote{ablasco@fas.harvard.edu}
    - KJ Boudreau
    - M Menietti
    - KR Lakhani
#institute: "Harvard's Institute for Quantitative Social Sciences"
date: "This version: `r format(Sys.Date(),'%B %d, %Y')`" 
bibliography: /Users/ablasco/Documents/papers.bib
toc: true
output: 
  beamer_presentation:
    template: /Users/ablasco/Templates/Pandoc/boyd.beamer
    slide_level: 2
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache=TRUE,
  echo=FALSE,
  message=FALSE,
  warning=TRUE,
  error=FALSE,
  tidy=TRUE,
  tidy.opts=list(width.cutoff=60),
  fig.width=9,
  fig.height=5,
  cache.path="Cache/",
  fig.path="Figures/")

# Prepare data
within(races, {
	year <- as.numeric(format(member_date, '%Y'))
	hours <- week1 + week2 + week3 + week4
	rating <- mm_rating
	submissions <- mm_events
	registrations <- mm_reg
	lpaid  <- log(paid)
	male <- gender=='Male'
	grad <- educ == "Doctorate/PhD" | educ == "Postgraduate/Master of arts"
	below30 <- age == "<20 years old" | age== "20-25 years old" | age =="26-30 years"
}) -> dat

```

Introduction
============

Contests and economic growth
-------------------------

Historically, awards offered by government

- navigation and cartography (Longitude awards in 1714)
- agricoltural innovation (Royal Agricultural Society awards 1900's)
- aviation industry (Orteig prize in 1919)
- architecture (Thomas Jefferson organized the first US design contest to build the White House in 1790)

Today, contests are common management tool

- Incentives for workers
- Philanthropic initiatives 
- Crowdsourcing internal activities to online communities of freelancers


The problem of contest design
-----------------------------

How to design a contest? Contest designers need to deal with:

- "Incentive" design problem 
	+ What is the optimal prize structure?

- "Competition" design problem
	+ a "race" competition? or in a "tournament" competition?


Race or tournaments?
====================


Examples by competition design
---------------------------

+ Races
	- Longitude prize (1714 and 2014)
	- Orteig Prize
	- Netflix prize
	- Ansari's X-prize

+ Tournaments
	- White House design contest
	- Golden Carrot Contest
	- DARPA Grand Challenges
	- X-Prize Challenges
	- European Commission's Horizon Challenge


Race or tournaments?
--------------------

- Contest designers face trade-off between "speed" and "quality"
	+ Example: seeking solutions to public health care problems (effective and timely)
	+ EC's Horizon Challenge (race) and 2014 Longitude Prize (tournament) about antibiotic resistance

- Efficiency concerns
	+ Fullerton & MacAfee (1992)'s result that only 2 competitors are optimal
	+ Prevent entry:
		+ **Minimum quality requirements**
		+ **Time deadlines**
	+ Example: online platforms running many contests

Prior literature
---------------

Theory: 

Patent races (e.g., Loury 1979, Nalebuff & Stiglitz 1983); contests (e.g., Lazear & Rosen 1981, Green Stokey 1983, Dixit 1987); debate on prizes or patents for R&D (Wright 1983); strategic equivalence (Baye Hoppe 2003)

Empirical works:

Non-experimental data on contests (e.g., Ehrenberg Bognanno 1990, Knoeber Thurman 1994, Eriksson 1999) and on races (Cockburn Henderson 1994); Laboratory experiments on contests and on races (see Dechenaux et al. 2014). 



In this paper
-------------

- We develop a model to examine contest designer's choice between races and tournaments in one framework
	+ The model extends Moldovanu and Sela (2001)'s static "all-pay" contest model

- We design and conduct a field experiment to examine predictions of the model
	+ An online community of highly skilled coders with experience in programming competitions


Model's basic setup
---------------

- "All-pay" contest with deadline $t_0$ and target $y_0$

- Cost: $C(a, q, t) = a^\alpha  y^\beta t^\gamma \qquad \text{with }\alpha,\gamma\leq-1, \beta\geq1$

- Payoff: $\pi_i = \sum_{k=1}^q p_k(y_i, t_i) v_k - C(a_i, y_i, t_i)$

- Equilibrium in a race: $t^*=b_{\text{race}}(a)$  and $y^*=y_0$

- Equilibrium in a tournament: $y^*=b_{\text{tournament}}(a)$  and $t^*=t_0$

Equilibrium performance
-----------------------

```{r}
integral1 <- function(v, a, n, alpha, a0) {
	v * a^(n-1-alpha) / (n-1-alpha) - 	v * a0^(n-1-alpha) / (n-1-alpha)
}
integral2 <- function(v, a, n, alpha, a0) {
	exponent <- (n-2-alpha)
	r <- ((n-2)*(exponent+1) - a*(n-1) * exponent ) / (exponent*(exponent+1))
	r0 <- ((n-2)*(exponent+1) - a*(n-1) * exponent ) / (exponent*(exponent+1))
	(1-v) * a^exponent * r - 	(1-v) * a0^exponent * r0
}
y <- function(v, a, n, alpha, beta) {
	V <- integral1(v, a, n, alpha, a0=0) + integral2(v, a, n, alpha, a0=0) 
	V^(1/beta)
}

#curve(y(v=1, a=x, n=4, alpha=-1, beta=2), n=1001, ann=FALSE)
#curve(y(v=0.75, a=x, n=4, alpha=-1, beta=2), lty=2, add=TRUE, n=1001)
#curve(y(v=0.5, a=x, n=4, alpha=-1, beta=2), lty=3, add=TRUE, n=1001)
#title("Equilibrium bids in a tournament")
#mtext("a ~ Uniform(0,1)", 1,3)
#mtext("y(a)", 2,3)

## RACE ############
y0 <- 0.15
payoff <- function(v, a, n, alpha, beta) {
 v * a^(n-1)  +
 (1-v) * (1-a) * a^(n-2) - 
 a^alpha * y0^beta  
}
abilityhat <- uniroot(payoff, interval=c(0,1), v=0.5, n=4, alpha=-1, beta=2)$root

curve(y(v=0.5, a=x, n=4, alpha=-1, beta=2), n=1001, ann=FALSE, lty=2)
segments(x0=abilityhat, y0=0, y1=y0, lty=3)
segments(x0=abilityhat, x1=1, y0=y0, y1=y0)
#title("Equilibrium bids in a race")
mtext("a ~ Uniform(0,1)", 1,3)
mtext("y(a)", 2,3)
legend("topright", c("race", "tournament"), lty=1:2, bty='n')
```

Behavioral predictions
----------------------

1. Competitors will enter a Tournament more than a Race
	+ Entry in a tournament is driven by "low-ability" competitors

4. On speed, Race  $\succ$ ("dominates") Tournament
	+ No clear dominance on performance

5. On performance, Tournament "with Reserve" $\succ$ Tournament and Race


Contest designer's problem
--------------------------

- Maximize expected revenues
- Revenues increase in performance, decrease in time of the winner ($y^w$, $t^w$)
- Consider $t_0$ and $y_0$ as given
- Expected payoff: $\pi_{cd} = E[y^w - \tau t^w  \mid y^w \geq y_0, t^w\leq t_0]$
- Main result:  if $\tau \geq \hat \tau$, the race should be preferred


The model
=========

Basic setup
------------

Imagine $i=1,...,n$ players competing for $k=1,...,q$ prizes of value $v_1\geq v_2\geq ...\geq v_q$ (normalized $\sum v_k=1$). 

Players simultaneously choose quality $y_i$ and time $t_i$ ($y_i/t_i$ speed). 

Each player has an ability $a_i$ drawn at random from a common cdf $F(\cdot)$ with pdf $f(\cdot)$.

The cost function $C(\cdot)$ is Cobb-Douglas: 

\begin{equation}
	C(a, q, t) = a^\alpha  y^\beta t^\gamma \qquad \text{with }\alpha,\gamma\leq-1, \beta\geq1
\end{equation}

- or denoting speed $(y/t)$ by $s$:

\begin{equation}
	C(a, q, t) = a^\alpha  y^{\beta^\prime} s^{\gamma^\prime} \qquad \beta^\prime=\beta+\gamma, \gamma^\prime=-\gamma. 
\end{equation}

Payoffs
--------
 
Player $i$'s payoffs:
\begin{equation}
	\pi_i = \sum_{k=1}^q p_k(y_i, t_i) v_k - C(a_i, y_i, t_i) 
\end{equation}

where $p_k(\cdot)$ is the prob. of winning prize $k$


Competition
-----------

- Let denote a deadline by $t_0$ and a minimum-quality target by $y_0$. 
 
 - We consider two competitive formats:
 
 	+ Race: competition with target where the first to achieve the target wins
 	+ Tournament: competition with deadline where the best wins


Probability
-----------

Let $y_{1:n}, ..., y_{n:n}$ denote the order statistics of the $y$'s.  Let denote the corresponding distribution functions by $F_{y_{1:n}},..., F_{y_{n:n}}$. 

Then the conditional probability of winning the first prize in a tournament  is 

$$
	\Pr(y_i\geq y_{n-1:n-1}) = F_{y_{n-1:n-1}} (y_i) = F(y_i)^{n-1}
$$

when $t_i \leq t_0$. And is zero otherwise.

$$
	\Pr(y_i xxxx ) =  [1 - F(y_i)]F(y_i)^{n-2}
$$

Probability 2
-----------

If $a\sim \text{Uniform}(0,1)$, then:

$$
	p_1(y) = y^{n-1}, \qquad p_2(y) =  [1 - y]y^{n-2}
$$

$$
	p_1(y)^\prime =  (n-1) y^{n-2}
$$

$$
	p_2(y)^\prime =  - y^{n-2} + (1-y)(n-2) y^{n-3} = y^{n-3} [(1-y)(n-1) - 1] 
$$

Contest designer's payoff
------------------------

Contest designer is risk neutral and wants to max quality while min time of the winner. 

$Z$ is the competition format. Let denote the race by $Z=1$ and the tournament by $Z=0$.
Let denote the winner's actions by $(y^w, t^w)$. 

The contest designer's expected payoff: 

\begin{equation}
	\pi_{cd} = E[ y^w - \tau t^w  \mid y^w \geq y_0, t^w\leq t_0, R]. 
\end{equation}

Solution concept 
-----------------

We solve the model for its unique symmetric Perfect Bayes Nash Equilibrium (the "equilibrium").

Let denote equilibrium bidding functions with respect to ability by $t(\cdot)$ and $y(\cdot)$. 

Consider Tournament first.

Maximization problem
-------------------

- Key observation: $t_i = t_0$ is a (weakly) dominant strategy

- This simplifies the maximization problem to:

\begin{equation}
	\max_{y} \hat\pi = \sum_{k=1}^q p_k(y) \hat v_k - a_i^\alpha  y^\beta 
\end{equation}

with $\hat v_k$ denoting each prize $v_k$ rescaled by a factor $t_0^\gamma$.

First order condition
----------------------

For each $i=1,..., n$, first order conditions are:

\begin{equation}
 \sum_{k=1}^q p_k^\prime(y) \hat v_k =  a_i^\alpha \beta y^{\beta-1}
\end{equation}


Solving differential equation
-----------------------------

Substituting the equilibrium function $y(\cdot)$ increasing in $a_i$ and with inverse $\phi(\cdot)$, together with a "change of variable" (moving $a_i=\phi(y_i)$ to the lhs): 

\begin{equation}
 \phi^{-\alpha}\sum_{k=1}^q \hat p^\prime_k(\phi) \phi^\prime v_k  = t_0^\gamma \beta y(a)^{\beta-1}
\end{equation}

Integrating both sides (using the "chain of derivatives" on the lhs):

\begin{equation}
	\sum_{k=1}^q \hat v_k  \int_{a_0}^{a} p^\prime_k(x) x^\alpha dx  + \beta y (a_0)^{\beta-1}
	=  \beta y (a)^{\beta-1} 
\end{equation}


Bidding function
-----------------

For every $i=1,..., n$:

- Time $t(a) = t_0$ 

- Equilibrium quality $y_i$ for competitor with ability $a$ is given by:

\begin{equation}
	y(a) = \left[y (a_0)^{\beta-1} + \frac 1\beta \sum_{k=1}^q \hat v_k  \int_{a_0}^{a} p^\prime_k(x) x^\alpha dx \right]^{1/{(\beta-1)}}
\end{equation}

with boundary condition $y(a_0)=0.$

Example
-------

If $a\sim \text{Uniform}(0,1)$ and $q=2$

First integral:

$$
	v_1 (n-1) \int_{0}^a x^{(n-2)-\alpha} dx = v_1 \frac{a^{(n-1) - \alpha}}{(n-1)-\alpha}
$$

Second integral:
$$
	v_2 \int_{0}^a x^{(n-3)-\alpha} [(1-x)(n-1) - 1]  dx 
$$

$$
	= v_2 \frac{a^{-\alpha +n-2} ((n-2) (-\alpha +n-1)- a (n-1) (-\alpha +n-2))}{(-\alpha +n-2) (-\alpha +n-1)}
$$

Impose normalization: $t_0^\gamma = 1$. 

Example 2
----------

```{r}
integral1 <- function(v, a, n, alpha, a0) {
	v * a^(n-1-alpha) / (n-1-alpha) - 	v * a0^(n-1-alpha) / (n-1-alpha)
}
integral2 <- function(v, a, n, alpha, a0) {
	exponent <- (n-2-alpha)
	r <- ((n-2)*(exponent+1) - a*(n-1) * exponent ) / (exponent*(exponent+1))
	r0 <- ((n-2)*(exponent+1) - a*(n-1) * exponent ) / (exponent*(exponent+1))
	(1-v) * a^exponent * r - 	(1-v) * a0^exponent * r0
}
y <- function(v, a, n, alpha, beta) {
	V <- integral1(v, a, n, alpha, a0=0) + integral2(v, a, n, alpha, a0=0) 
	V^(1/beta)
}

curve(y(v=1, a=x, n=4, alpha=-1, beta=2), n=1001, ann=FALSE)
curve(y(v=0.75, a=x, n=4, alpha=-1, beta=2), lty=2, add=TRUE, n=1001)
curve(y(v=0.5, a=x, n=4, alpha=-1, beta=2), lty=3, add=TRUE, n=1001)
title("Equilibrium bids in a tournament")
mtext("a ~ Uniform(0,1)", 1,3)
mtext("y(a)", 2,3)
```


Bidding function in a race
-------------------------

For every $i=1,..., n$ with $a_i \geq \hat a$

- Quality $y(a) = y_0$ 

- Time:

\begin{equation}
	t(a) = \left[t (t_0)^{\gamma-1} + \frac 1\gamma \sum_{k=1}^q \tilde v_k  \int_{a_0}^{a} \hat p^\prime_k(x) x^\alpha dx \right]^{1/{(\gamma-1)}}
\end{equation}

with $\tilde v_k = v_k / y_0^{\beta}$. 

Otherwise, when $a_i < \hat a$, and $y(a) < y_0$. 

Zero profit
------------

The zero profit condition for the marginal player: 

\begin{equation}
	\sum p_k (y_0, t_0) v_k = \hat a^\alpha y_0^\beta t_0^\gamma
\end{equation}

Hence, the marginal ability is pinned down:

\begin{equation}
	\hat a = \left[\sum p_k (y_0, t_0) v_k / y_0^\beta t_0^\gamma\right]^{1/\alpha}
\end{equation}


Example 2
-----------

If $a\sim \text{Uniform}(0,1)$, then ZPC

$$
	p_1(y) = y^{n-1}, \qquad p_2(y) =  [1 - y]y^{n-2}
$$

($t_0^\gamma=1$)

$$
	\pi_i = v_1  p_1(a)  + v_2 p_2(a) - a^\alpha y_0^\beta
$$

$$
	(ZPC) \qquad v_1  a^{n-1}  + (1-v_1)[1 - a]a^{n-2} - a^\alpha y_0^\beta = 0
$$


Example
----------

```{r}
y0 <- 0.15
payoff <- function(v, a, n, alpha, beta) {
 v * a^(n-1)  +
 (1-v) * (1-a) * a^(n-2) - 
 a^alpha * y0^beta  
}
abilityhat <- uniroot(payoff, interval=c(0,1), v=0.5, n=4, alpha=-1, beta=2)$root
# $
curve(y(v=0.5, a=x, n=4, alpha=-1, beta=2), n=1001, ann=FALSE, lty=2)
segments(x0=abilityhat, y0=0, y1=y0, lty=3)
segments(x0=abilityhat, x1=1, y0=y0, y1=y0)
title("Equilibrium bids in a race")
mtext("a ~ Uniform(0,1)", 1,3)
mtext("y(a)", 2,3)
```


```{r, include=FALSE}
# TIME ... 
integral1 <- function(v, a, n, alpha) {
	v * a^(n-1-alpha) / (n-1-alpha)
}
integral2 <- function(v, a, n, alpha) {
	exponent <- (n-2-alpha)
	numer <- (n-2)*(exponent+1) - a*(n-1) * exponent
	denom <- (exponent*(exponent+1))
	(1-v) * a^exponent * numer /  denom
}
race <- function(v, a, n, alpha, beta, gamma, a0) {
	v <- v / y0^beta
	int1 <- integral1(v, a, n, alpha) - integral1(v, a0, n, alpha)
	int2 <- integral2(v, a, n, alpha) - integral2(v, a0, n, alpha)
	y <- int1 + int2
	( 1 +  y / gamma)^(1/(gamma-1))
}
payoff <- function(v, a, n, alpha, beta) {
 v * a^(n-1)  +
 (1-v) * (1-a) * a^(n-2) - 
 a^alpha * y0^beta  
}
y0 <- 0.01
#abilityhat <- uniroot(payoff, interval=c(0,1), v=0.5, n=4, alpha=-1, beta=2)
#curve(race(v=0.5, a=x, n=4, alpha=-1, gamma=-2, beta=2, a0=abilityhat), n=1001, ann=FALSE, lty=2)
```

Experimental design
==================

The context: Topcoder.com
----------

Recruit 229 competitors on Topcoder for eight-day programming competition

Three key factors: 

- Platform members are sophisticated competitors
- Observe measures of skills
- Rich data analytics about performance and timing

The contest
--------

- Total prize purse $41000
	+ Grand prizes of $6000 across competition styles
	+ Room prizes of $1000 and $100 for 1st and 2nd
- Task solving Named Entity Recognition Problem in medical research


Experimental design
-------------------

- 8 day submission phase
- Split into 24 rooms of 10 and 15 competitors
- 3x2 experimental design (Race, Tournament, Reserve) x (10, 15)

Data
-----


```{r, results='asis'}
descriptives <- function(dat, treat=races$treatment) {
	balance.test <- function(x, treat=races$treatment) {
		if (is.logical(x)) return(fisher.test(table(x, treat))$p.val)
		l <- split(x, treat)
		kruskal.test(l)$p.val
	}
	mu <-sapply(dat, mean, na.rm=TRUE)
	q50 <-sapply(dat, median, na.rm=TRUE)
	lo <-sapply(dat, min, na.rm=TRUE)
	hi <-sapply(dat, max, na.rm=TRUE)
	std <-sapply(dat, sd, na.rm=TRUE)
	pval <- sapply(dat, balance.test)
	n <- sapply(dat, function(x) sum(!is.na(x)))
	tab <- cbind(mu, q50, std, lo, hi, n, pval)
	colnames(tab) <- c("Mean", "Median","St.Dev.", "Min", "Max", "Obs.", "P-value")
	return(tab)
}
covars <- with(dat, data.frame(year, rating, registrations, submissions, lpaid, nwins, ntop10, risk, hours, male, timezone, grad, below30))
tab <- descriptives(covars)
xtab <- xtable(tab)
digits(xtab) <- c(1, 1, rep(0, ncol(tab)-2), 3)
align(xtab) <- c("@{}l", rep("r", ncol(tab)))
print(xtab, comment=FALSE)
```

Skill rating distribution
-------------------------

```{r}
attach(races)
rated <- !is.na(mm_rating)
rating_l <- split(mm_rating[rated], treatment[rated])
rating_l.pdf <- lapply(rating_l, density, from=500, bw='nrd', kernel='gaussian')
rating_l.test <- kruskal.test(rating_l)

plot.density <- function(list, ...) {
	xlim <- range(sapply(list, function(x) range(x$x)))
	ylim <- range(sapply(list, function(x) range(x$y)))
	legend.names <- capitalize(names(list))
	plot.new()
	plot.window(xlim=xlim, ylim=ylim, ...)
	sapply(1:3, function(i) lines(list[[i]], lty=i, lwd=2))
	legend("topright", bty='n', legend.names, lty=1:3)
	xseq <- pretty(seq(xlim[1], xlim[2], length=20))
	axis(1, at=xseq, labels=xseq)
}

plot.density(rating_l.pdf)
detach(races)
```


Results
========

Distribution of room outcomes
--------------

```{r, fig.width=9, fig.height=4}

# CREATE DATA for ROOMS
entrants <- aggregate(submit ~ room + room_size + treatment, data=races, sum)
cap.scores <- function(x, baseline=0.792867) ifelse(x<baseline, baseline, x)
races$final.cap <- cap.scores(races$final/1e6)
final <- aggregate(final.cap ~ room + room_size + treatment, data=races, FUN=mean, na.rm=TRUE)
firstsub <- as.numeric(with(races, difftime(timestamp, "2015-03-08 12:00:00 EDT", units='hours')))
speed <- aggregate(firstsub ~ room + room_size + treatment, data=races, FUN=mean, na.rm=TRUE)


par(mfrow=c(1, 3))
boxplot(submit ~ treatment, data=entrants, xlab='Room entrants'
	, col="lightgray", horizontal=TRUE, boxwex=0.5); title("Participation")
boxplot(final.cap ~ treatment, data=final, xlab='Room final score (avg.)'
	, col="lightgray", horizontal=TRUE, boxwex=0.5); title("Performance")
boxplot(firstsub ~ treatment, data=speed, xlab='Time first submission (hours)'
	, col="lightgray", horizontal=TRUE, boxwex=0.5); title("Speed")
```


Greater participation in the Tournament
---------------------------------------

\columnsbegin
\column{.3\textwidth}

```{r}
boxplot(submit ~ treatment, data=entrants, xlab='Room entrants'
	, col="lightgray", horizontal=TRUE, boxwex=0.5); title("Participation")
```

\column{.7\textwidth}

```{r}
t.test(submit ~ treatment!='tournament', alternative='greater', var.equal=FALSE, data=entrants)
```
\columnsend


No evidence of skill-based selection
------------------------------------

```{r skill-based-selection, fig.width=9, fig.height=7}
covars <- covars[dat$submit, ]
covars$nwins <- covars$nwins>0
covars$ntop10 <- covars$ntop10>0
covars$submission_pc <- covars$submissions  / covars$registrations
plot.covars <- function(x, ...) {
	stars.pval <- function(x) {
		if (x <= 0.01) return("***")
		if (x <= 0.05) return("** ")
		if (x <= 0.1) return("*  ")
		else return("   ")
	}
#	l <- dat$treatment[dat$submit]
	l <- ifelse(dat$treatment[dat$submit]=='tournament', "Tourn", "Other")
	if (!is.logical(x)) {
		x.l <- split(x, l)
#		p <- kruskal.test(x.l)$p.value
		p <- t.test(x ~ l)$p.value
		pval <- sprintf("P-value=%0.3f%s", p, stars.pval(p))
		boxplot(x ~ l, horizontal=TRUE, las=2, xlab=pval, ...)
	} else {
		tab <- table(x, l)
		p <- fisher.test(tab)$p.value
		pval <- sprintf("P-value=%0.3f%s", p, stars.pval(p))
		barplot(prop.table(tab, 2), horiz=TRUE, las=2, xlab=pval, ...)
	}
	return(p)
}
par(mfrow=c(4,4), mar=c(4,4,3,2))
pval <- sapply(1:ncol(covars), function(i) plot.covars(covars[, i], main=colnames(covars)[i]))
#pval #p.adjust(pval, method='bonferroni') # only one remains
```

Deal with noise in performance
------------------------------

```{r performance-appendix, fig.width=9, fig.cap="Scores over time\\label{scores over time}"}
data(scores)

## Order scores by time
scores <- scores[order(scores$submission), ]
index <- tapply(1:nrow(scores), scores$coder_id, tail, 1) # last submission
scores <- scores[index, ]
scores <- scores[order(scores$timestamp), ] # order by time
scores$final[is.na(scores$final)] <- 0  # Impute 2 missing scores

## https://en.wikipedia.org/wiki/Winsorized_mean
cap.scores <- function(x, threshold) ifelse(x<threshold, threshold, x)
cap.01 <- quantile(scores$final, na.rm=TRUE, p=0.1)
scores$final.cap <- cap.scores(scores$final, threshold=cap.01) 
scores$final.cap2 <- cap.scores(scores$final, threshold=792867)  # Baseline

par(mfrow=c(3, 1), mar=c(2,3,3,1))
plot(final/1e6 ~ timestamp, data=scores
	, type='l', col='lightgray', ylab="Final score")
title("Observed")
abline(h=0.792867, lty=3) # Baseline
abline(h=0.817866, lty=2, col=2) # Target

plot(final.cap/1e6 ~ timestamp, data=scores
	, type='l', col='lightgray', ylab="Final score")
title("Winsorized at 10 percent")
abline(h=0.792867, lty=3) # Baseline
abline(h=0.817866, lty=2, col=2) # Target

plot(final.cap2/1e6 ~ timestamp, data=scores
	, type='l', col='lightgray', ylab="Final score")
title("Capped at the baseline")
abline(h=0.792867, lty=3) # Baseline
abline(h=0.817866, lty=2, col=2) # Target


# Merge with treatment
#z <- merge(scores, races[, c("coder_id", "room", "treatment")], by='coder_id')
#z.mean <- aggregate(final.cap ~ room + treatment, data=z, mean)
#z.mean2 <- aggregate(final.cap2 ~ room + treatment, data=z, mean)
#par(mfrow=c(2, 1), mar=c(4,5, 2,2))
#boxplot(final.cap ~ treatment, data=z.mean, horizontal=TRUE, las=2)
#boxplot(final.cap2 ~ treatment, data=z.mean2, horizontal=TRUE, las=2)
```

No evidence of higher performance in Reserve
---------------------------------------------------------

\columnsbegin
\column{.3\textwidth}

```{r}
boxplot(final.cap ~ treatment, data=final, xlab='Room final score (avg.)'
	, col="lightgray", horizontal=TRUE, boxwex=0.5); title("Performance")
```

\column{.7\textwidth}

```{r}
t.test(final.cap ~ treatment=='reserve', var.equal=FALSE, data=final) # p = 0.94
```
\columnsend



No evidence of difference in performance between Tournament and Race
---------------------------------------------------------

\columnsbegin
\column{.3\textwidth}

```{r}
boxplot(final.cap ~ treatment, data=final, xlab='Room final score (avg.)'
	, col="lightgray", horizontal=TRUE, boxwex=0.5); title("Performance")
```

\column{.7\textwidth}

```{r}
t.test(final.cap ~ treatment, var.equal=FALSE, data=final, subset=treatment!='reserve') # p = 0.94
```
\columnsend


Speed was higher in the Race 
-----------------------------

\columnsbegin
\column{.3\textwidth}

```{r}
boxplot(firstsub ~ treatment, data=speed, xlab='Time first submission (hours)'
	, col="lightgray", horizontal=TRUE, boxwex=0.5); title("Speed")
```

\column{.7\textwidth}

```{r}
t.test(firstsub ~ treatment=='race', alternative='greater', var.equal=FALSE, data=speed)  # p = 0.03
```
\columnsend

To summarize
------------

1. Participation was higher in the Tournament
	+ It was not driven by low-skill competitors

2. No evidence of a difference in performance

3. Competitors worked faster in a race

Interpretation: 

- Taken together, though competitors seemed to "like" tournaments more, they "worked" more in a race.
- Tournament with reserve does not seem to be better than tournament


Structural approach
===================

Modeling individual behavior
----------------------------

Identification of causal effect of competition on individual behavior is problematic

1. Actions are correlated, violating one key assumption of Rubin's potential outcomes causality model
2. Censoring (entry/exit decisions are only partially observed)
3. Dynamics

Under our model, however, we have:

$$y = 1 \iff \text{ability} > a_0$$ 

$\leadsto$ "single-index" models have nice structural interpretation.

Entry decision
--------------

```{r, echo=FALSE, include=FALSE}
# Create dataset for regression imputing missing values at random
within(dat, {
	set.seed(2159)
#	set.seed(25978)
	rating.100 <- rating / 100
	hours.imp <- impute(hours, "random")
	risk.imp <- impute(risk, "random")
	grad.imp <- impute(grad, "random")
	male.imp <- impute(male, "random")
	below30.imp <- impute(below30, "random")
	timezone.imp <- impute(timezone, "random")
	expert <- cut(submissions, quantile(submissions[!is.na(rating)]), include=TRUE)
}) -> dat.imp


# First model has no covariates (i.e., $\gamma=0$)
summary(m0 <- glm(submit ~ treatment, binomial(logit), data=dat.imp))

# The second model adds the main skill rating measure which is available for 2/3 of our population. It seemed better to rescale rating in 100-point units and center it on the median value. Thus, the estimate of the intercept can be easily transformed in the probability of participation of the median rated individual assigned to a room with a race competition style.

summary(m1 <- update(m0, " ~ . + rating.100"))

# The third column adds time availability (hours)
summary(m2 <- update(m1, " ~ . + hours.imp"))

# ... add demographics to m2
summary(m3 <- update(m2, " ~ . + timezone.imp + grad.imp + below30.imp + male.imp + risk.imp"))
```

```{r, results='asis'}
# Compare models 
models <- list(m0, m1, m2, m3)
stargazer(models, digits=2, header=FALSE, single.row = FALSE
	, no.space=TRUE, float=FALSE, column.sep.width='1pt'
	, font.size='scriptsize', style='io'
	, dep.var.labels.include=FALSE)
```

Model's fit is ok
------------

```{r, fig.width=9}
summarize.fit <- function(x) {
	yhat <- predict(x)
	with(x, plot(jitter(y) ~ yhat, col=ifelse(yhat>0,'brown', 'blue'), pch=16, xlab="ability ~ Logistic"))
	curve(ilogit, add=T)
}
par(mfrow=c(1, 2))
summarize.fit(m1); title("M1")
summarize.fit(m3); title("M3")
```

Entry decision across competition styles
----------------------------------------

```{r, include=FALSE}
m.selected <- step(m3)
# Using Akaike criterion to select the best model, we now compute the model for each treatment alone
summary(m.race <- update(m.selected, "~ . -treatment", subset=treatment=='race'))
summary(m.tour <- update(m.selected, "~ . -treatment", subset=treatment=='tournament'))
summary(m.rese <- update(m.selected, "~ . -treatment", subset=treatment=='reserve'))
```

```{r, results='asis'}
# Compare models
models <- list(m.selected, m.race, m.tour, m.rese)
stargazer(models, digits=2, header=FALSE, single.row = FALSE
	, no.space=TRUE, float=FALSE, column.sep.width='1pt'
	, font.size='scriptsize', style='io'
	, dep.var.labels.include=FALSE
	, column.labels=c("All", "Race", "Tourn", "Rese")
	, model.numbers=FALSE)
```

Model's fit
------------

```{r, fig.width=9}
summarize.fit <- function(x) {
	yhat <- predict(x)
	with(x, plot(jitter(y) ~ yhat, col=ifelse(yhat>0,'brown', 'blue'), pch=16, xlab="ability ~ Logistic"))
	curve(ilogit, add=T)
}
par(mfrow=c(1, 3))
summarize.fit(m.race); title("Race")
summarize.fit(m.tour); title("Tournament")
summarize.fit(m.rese); title("Reserve")
```

Incorporating Scores
--------------------

Individual scores are censored $\leadsto$ OLS is problematic. 

We examine "production speed" $y_i$ ($=\text{score}_i/t_i$) at a given point in time. 

Then, our data's likelihood is (e.g., Tobit):

$$
	\mathcal{L} = \prod_{i=1}^N \Pr(Y \geq 0)^{1-I(y_i)} \times
		\Pr(Y=y_i)^{I(y_i)}. 
$$

Under the model's equilibrium, this becomes

$$
	\mathcal{L} = \prod_{i=1}^N [1-F(a_{0,i})]^{1-I(y_i)} \times
		f(b(\text{ability}_i)=y_i)^{I(y_i)}. 
$$


Estimation
----------

- We use parametric $F$ known up to a vector $\theta$, that we estimate from the data.

- Compare against Tobit model (our main benchmark)

- Alternatively, replace $F$ with skill rating's $\hat F$ (our second benchmark)


Benchmark
--------------

```{r, include=FALSE}
library("AER")
dat.imp$final0 <- impute(dat.imp$final/1e6, "zero")
fit.tobit <- tobit(final0 ~ treatment, left=0.1, data=dat.imp)
fit.tobit2 <- tobit(final0 ~ treatment + rating.100 + hours.imp, left=0.1, data=dat.imp)
ols <- glm(final0 ~ treatment, data=dat.imp)
ols2 <- glm(final0 ~ treatment + rating.100 + hours.imp, data=dat.imp)
```

```{r, results='asis'}
# Compare models
models <- list(fit.tobit, fit.tobit2, ols, ols2)
stargazer(models, digits=2, header=FALSE, single.row = FALSE
	, no.space=TRUE, float=FALSE, column.sep.width='1pt'
	, font.size='scriptsize', style='io'
	, dep.var.labels.include=FALSE
	, keep.stat=c('n','ll'))
```


Estimated probability of entry
-------------------------------

```{r}
params <- coef(fit.tobit)
curve(pnorm(x, mean=params[1]), ylab='Prob. of entry', xlab="Ability", from=0, to=4)
curve(pnorm(x, mean=params[1] + params[2]), add=TRUE, lty=2)
curve(pnorm(x, mean=params[1] + params[3]), add=TRUE, lty=3)
legend("bottomright", bty='n', levels(races$treatment), lty=1:3)
```



Appendix
========

References {.allowframebreaks}
-----------
