---
title: "Races or Tournaments?"
thanks: "Blasco: Harvard Institute for Quantitative Social Science, Harvard University, 1737 Cambridge Street, Cambridge, MA 02138 (email: ablasco@fas.harvard.edu)"
subtitle: "[DRAFT: Please, do not distribute]"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
author: 
    - Andrea Blasco
    - Kevin J. Boudreau
    - Karim R. Lakhani
    - Michael Menietti
---

\begin{abstract}

\noindent A wide range of economic and social situations are decided by either a
race or a tournament. In such contests, agents choose whether and how
much to exert some costly effort to increase the probability of being
awarded a prize under uncertainty about the other agents types or
actions. In theory, whenever the sponsor of the competition prefers
competitors' performance over the time to complete a particular task,
the expected outcomes of a tournament setup should be either equal or
greater than those of a race. Yet, a race might be more efficient from
an economic point of view as it may prevent unnecessary costs due to an
excess of participation. We examine this trade-off empirically. We
report the results of a field experiment conducted on a leading
crowdsourcing platform where we compare the outcomes (efforts, quality,
and diversity of outputs) of three alternative competitive situations
motivated by theory: the race, the tournament, and the tournament with a
quality requirement.

\smallskip\noindent 
JEL Classification: XX; XX; XX;

\smallskip\noindent 
Keywords: xx; xx; xx;
\end{abstract}
\clearpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache=TRUE,
  echo=FALSE,
  message=FALSE,
  warning=TRUE,
  error=FALSE,
  tidy=FALSE,
  fig.width=5,
  fig.height=5,
  cache.path="Cache/",
  fig.path="Figures/")
```


```{r child = '1_intro.Rmd'}
```

```{r child = '2_model.Rmd'}
```

```{r child = '4_analysis.Rmd'}
```

```{r child = '5_structural.Rmd'}
```

\appendix

```{r child = 'appendix.Rmd'}
```


References
=============
