---
title: "Races or Tournaments?"

thanks: "Blasco: Harvard Institute for Quantitative Social Science, Harvard University, 1737 Cambridge Street, Cambridge, MA 02138 (email: ablasco@fas.harvard.edu)."

date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"

author: 
    - Andrea Blasco
    - Kevin J. Boudreau
    - Karim R. Lakhani
    - Michael Menietti

abstract: "A wide range of economic and social situations are decided by either a race or a tournament. In either situations, agents choose whether and how much to exert some costly effort to increase the probability of being awarded a prize under the uncertainty about the actions of other agents. While a tournament yield outcomes greater than those of a race, the latter prevents unnecessary costs due to an excess of participation. We examine this trade-off empirically. We report the results of a field experiment conducted online where we compare the outcomes --- efforts, quality, and diversity of outputs --- of three alternative competitive situations motivated by theory: the race, the tournament, and the tournament with a quality requirement."

JEL: 
  - xxx
  - xxx
  - xxx

keywords: 
  - xxxx
  - xxxx xxxx

amsthm:
  plain:
  - theorem
  - proposition
  - corollary
  - definition

fontsize: 12pt

...


```{r, echo=FALSE, warnings=FALSE, message=FALSE}

#########################
load(".RData")
set.seed(4881) # for bootstrap
require(magrittr)
require(xtable)
require(stargazer)
#########################

knitr::opts_chunk$set(
  cache=FALSE,
  echo=FALSE,
  message=FALSE,
  warning=TRUE,
  error=FALSE,
  tidy=FALSE,
  cache.path="Cache/",
  fig.path="Figures/")
```

```{r child = '1_intro.Rmd'}
```

```{r child = '2_methods.Rmd'}
```

```{r child = '3_results.Rmd'}
```

```{r child = '4_discussion.Rmd'}
```


References
=============
