---
title: "Races or Tournaments? [PRELIMINARY AND INCOMPLETE]"

thanks: "Blasco: Harvard Institute for Quantitative Social Science, Harvard University, 1737 Cambridge Street, Cambridge, MA 02138 (email: ablasco@fas.harvard.edu)."

date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"

author: 
    - Andrea Blasco
    - Kevin J. Boudreau
    - Karim R. Lakhani
    - Michael Menietti

abstract: "We examine the performance of two different choices of contest design: the race (where the winner is the first to achieve a minimum quality) and the tournament (where the winner is the one with the highest quality in a given period). After characterizing  the optimal design, we report results of a field experiment conducted to compare the performance of three alternatives motivated by theory: the race, the tournament, and the tournament with a minimum quality requirement. Outcomes in a race are of comparable quality, supplied faster, and with lower participation rates. Based on these findings, we show the optimal design under several counterfactual situations."

JEL: 
  - xxx
  - xxx
  - xxx

keywords: 
  - xxxx
  - xxxx xxxx

fontsize: 12pt
...

```{r setup, echo=FALSE}
load(".RData")

set.seed(4881)

require(knitr)
require(magrittr)
require(xtable)
require(stargazer)

options(digits=3, xtable.caption.placement='top', xtable.comment=FALSE)

opts_chunk$set(
	cache=TRUE,
	echo=TRUE,
	message=TRUE,
	warning=TRUE,
	error=TRUE,
	tidy.opts=list(width.cutoff=60),
	tidy=TRUE,
	cache.path="Cache/",
	fig.path="Figures/")
```

```{r child = '1_intro.Rmd'}
```

```{r child = '2_model.Rmd'}
```

```{r child = '3_data.Rmd'}
```

```{r child = '4_results.Rmd'}
```

```{r child = '5_discussion.Rmd'}
```
 
```{r session_info}	
devtools::session_info()
```

References
=============
