---
title: "Races or Tournaments? [PRELIMINARY AND INCOMPLETE]"

thanks: "Blasco: Harvard Institute for Quantitative Social Science, Harvard University, 1737 Cambridge Street, Cambridge, MA 02138 (email: ablasco@fas.harvard.edu)."

date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"

author: 
    - Andrea Blasco
    - Kevin J. Boudreau
    - Karim R. Lakhani
    - Michael Menietti

abstract: "A wide range of economic and social situations are decided by either a race or a tournament. In either situations, agents choose whether and how much to exert some costly effort to increase the probability of being awarded a prize under the uncertainty about the actions of other agents. While a tournament yield outcomes greater than those of a race, the latter prevents unnecessary costs due to an excess of participation. We examine this trade-off empirically. We report the results of a field experiment conducted online where we compare the outcomes --- efforts, quality, and diversity of outputs --- of three alternative competitive situations motivated by theory: the race, the tournament, and the tournament with a quality requirement."

JEL: 
  - xxx
  - xxx
  - xxx

keywords: 
  - xxxx
  - xxxx xxxx

amsthm:
  plain:
  - theorem
  - proposition
  - corollary
  - definition

fontsize: 12pt

...


```{r, echo=FALSE, warnings=FALSE, message=FALSE}

#########################
load(".RData")
set.seed(4881) # for bootstrap
require(magrittr)
require(xtable)
require(stargazer)
#########################

knitr::opts_chunk$set(
  cache=FALSE,
  echo=FALSE,
  message=FALSE,
  warning=TRUE,
  error=FALSE,
  tidy=FALSE,
  cache.path="Cache/",
  fig.path="Figures/")

# Helper functions
render.table <- function(x, caption, label, add=NULL, align=NULL, digits=1, notes=NULL) {
  if (is.null(align)) {
    align <- c("@{}l", rep("c", ncol(x)))
  }
  table.head <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]\n"
  table.mid <- "\\hline \\\\[-1.86ex]\n"
  table.bottom <- table.head
  add.to.cmd <- c(table.head, table.mid, table.bottom)
  add.to.row <- list(pos = list(-1, 0, nrow(x))
                , command = add.to.cmd)
  if (!is.null(add)) {
    add.to.row$command  <- c(add.to.row$command, add$cmd)
    add.to.row$pos[[4]] <- add$pos
  }
  cat("\\begin{table}\n")         
  cat("\\centering\n")                  
  cat(sprintf("\\caption{%s}\n", caption))
  cat(sprintf("\\label{%s}\n", label))
  print(xtable(x, caption, label, align, digits)
      , add.to.row=add.to.row
      , hline.after=NULL
      , floating=FALSE
      , comment=FALSE)
  if (!is.null(notes)) {
    cat("\\begin{tablenotes}\n")                  
    cat(sprintf("%s\n", notes))
    cat("\\end{tablenotes}\n")
  }
  cat("\\end{table}\n")   
}          




```

```{r child = '1_intro.Rmd'}
```

```{r child = '2_model.Rmd'}
```

```{r child = '3_data.Rmd'}
```

```{r child = '4_results.Rmd'}
```

```{r child = '5_discussion.Rmd'}
```


References
=============
