---
title: "Races or Tournaments? Theory and Evidence from the Field"

subtitle: "[PRELIMINARY AND INCOMPLETE]"

thanks: "Blasco: Harvard Institute for Quantitative Social Science, Harvard University, 1737 Cambridge Street, Cambridge, MA 02138 (email: ablasco@fas.harvard.edu)."

date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"

author: 
    - Andrea Blasco
    - Kevin J. Boudreau
    - Karim R. Lakhani
    - Michael Menietti

abstract: "We examine the performance of two different choices of contest design: the race (where the winner is the first to achieve a minimum quality) and the tournament (where the winner is the one with the highest quality in a given period). After characterizing  the optimal design, we report results of a field experiment conducted to compare the performance of three alternatives motivated by theory: the race, the tournament, and the tournament with a minimum quality requirement. Outcomes in a race are of comparable quality, supplied faster, and with lower participation rates. Based on these findings, we show the optimal design under several counterfactual situations."

JEL: 
  - M15 # IT Management
  - M52 # Personnel Economics, compensations and compensation methods
  - O31 # Innovation and Invention: Processes and Incentives

keywords: 
  - races
  - tournaments
  - contest theory
  - crowdsourcing
  - innovation

fontsize: 11pt
setspace: \doublespacing

output:
  pdf_document:
    toc: false
    keep_tex: false
    fig_crop: false
    fig_width: 5
    fig_height: 5
    fig_caption: true
    highlight: tango #kate
    template: ~/Templates/Pandoc/paper.latex
    includes:
      in_header: Templates/header.tex
#    before_body:   Templates/doc_prefix.tex
#    after_body:    Templates/doc_suffix.tex
    pandoc_args: [
      "--bibliography", "library.bib" #, "--natbib"
    ]
...

```{r setup, echo=FALSE, message=FALSE}

knitr::opts_chunk$set(
	cache=TRUE,
	echo=FALSE,
	message=FALSE,
	warning=FALSE,
	error=TRUE,
	tidy=FALSE,
	tidy.opts=list(width.cutoff=60),
	cache.path="Cache/",
	fig.path="Figures/")

set.seed(4881)

require(races)
require(magrittr)
require(xtable)
require(stargazer)

# Helper function
percent  <- function(x, digits=0, ...) round(100 * x, digits, ...)
difftime2 <- function(...) as.numeric(difftime(...))

options(digits=3, xtable.caption.placement='top', xtable.comment=FALSE)

# Prepare data
within(races, {
	year <- as.numeric(format(member_date, '%Y'))
	hours <- week1 + week2 + week3 + week4
	rating <- mm_rating
	submissions <- mm_events
	registrations <- mm_reg
	lpaid  <- log(paid)
	male <- gender=='Male'
	grad <- educ == "Doctorate/PhD" | educ == "Postgraduate/Master of arts"
	below30 <- age == "<20 years old" | age== "20-25 years old" | age =="26-30 years"
}) -> dat

# Documents
child_docs <- c("1_intro.Rmd", "2_model.Rmd", "3_data.Rmd", "4_results.Rmd", "5_discussion.Rmd")	
```

```{r child = child_docs}
```
 
```{r session_info}	
devtools::session_info()
```

References
=============
